---
title: "GW_Flux_Density_switched"
author: "Casey McGrath"
date: "November 7, 2018"
output: html_document
---

```{r setup, include=FALSE}
library(xts)
library(graphics)
library(dygraphs)
library(lubridate)
library(magrittr)

gr <- (1+sqrt(5))/2 # golden ratio, for figures
# The following chunk options can be overridden in chunk headers
require("knitr")
require("xtable")
opts_chunk$set( # a function in the knitr package
  echo = TRUE,  # show code chunks
  fig.asp=1/gr, # <---- set default figure aspect ratio
  fig.align="center")
```

**Calculating Gas Well Flux Density for Lyon Deep Soil Warming Project**

To calculate the gas well flux density for the gas wells across the hillslope the following calculations are used:

The first step is to calculate the molar density in $umol/m^3$ by using the molar density and volumetric water content (VWC). This can be done using the ideal gas law:

$$
PV= nRT
$$
Where $P$ is pressure (atm) , $V$ is volume (mL), $n$ is the moles of CO2, $R$ is the known gas constant ($J/mol$) and $T$ is the temperature (K). Here we import the datasets for temperature ($T$) and the Pressure ($P$) and the gas constant ($R$) are constants and moles of C02/ volume  is the variable in question: 

```{r loading and molar_density, echo=FALSE}

#set working directory
setwd("C:\\Users\\14014\\Desktop\\UH_Soil_Warming\\Gas_wells\\R_input")
#constants GWdata$airmd
P <- 0.9725935829 *101.325#pressure from local weather station
R <- 8.314 #gas constant (J/mol/K)

#bring in CO2 data and assign headers
GWdata <-read.csv(file="GW_2019_10_18.csv", skip=1, header=FALSE, stringsAsFactors = F) 
names(GWdata)<-c("location", "depth", "well1", "well2", "well3", "Temp", "VWC", "Bd")

#bring in temp and VWC data and assign headers



GWdata$airmd <- P*1000/(R*(GWdata$Temp+273.15)) # find the molar density of air at the time of sampling for each temp

GWdata$CO2mdx1 <- GWdata$airmd*(GWdata$well1/1000000) #molar density of well 1 shallow 

GWdata$CO2mdx2 <- GWdata$airmd*(GWdata$well2/1000000) #molar density of well 2 middle 

GWdata$CO2mdx3 <- GWdata$airmd*(GWdata$well3/1000000) #molar denstiy of well 3 deep
```

Next we take the derivative of the CO2 molar density vs. depth for dc/dz (C2-C1/Z2-Z1)

```{r derivative}
n <- length(GWdata$CO2mdx1)

#shallow to middle gas well derivative for loop 
GWdata$deriv1 <- numeric(n)
for (i in 1:n){
    if (GWdata$CO2mdx2[i] > GWdata$CO2mdx1[i]){
      GWdata$deriv1[i] <- (GWdata$CO2mdx2[i] - GWdata$CO2mdx1[i])/0.2 
    }else { 
      GWdata$deriv1[i] <- (GWdata$CO2mdx1[i]-GWdata$CO2mdx2[i])/0.2 
    }
}

#middle to deep gas well derivative for loop
GWdata$deriv2 <- numeric(n)
for (i in 1:n){
    if (GWdata$CO2mdx3[i] > GWdata$CO2mdx2[i]){
      GWdata$deriv2[i] <- (GWdata$CO2mdx3[i] - GWdata$CO2mdx2[i])/0.2 
    }else { 
      GWdata$deriv2[i] <- (GWdata$CO2mdx2[i]-GWdata$CO2mdx3[i])/0.2 
    }
}


```

Then we find the diffusion coefficents for each depth section from bulk density (Bd), water content (VWC) and temperature (Temp). 

```{r coefficents}
####constants####

#diffusion coefficent of air (mn^2/s^-1)
Doa <- 15.7 #not converted to m for later umol cancellation 

#particle denstiy 
Pd <- 2.65

####coefficent calculations #### 

#total porosity = 1 - (Bulk density/Particle density) * VWC 
GWdata$Tp<- (1-(GWdata$Bd/Pd))- GWdata$VWC 

#tourtisity = 0.95 * Total porosity ^ 1.95 
GWdata$epsilon <- 0.95*(GWdata$Tp^1.95)

#diffusion coefficent Do
GWdata$Do <- Doa * ((GWdata$Temp+273.15)/293.15)^1.75 * (101.3/P)

# #diffusion coefficent D
GWdata$D <- GWdata$Do * GWdata$epsilon 
```

Now we get the flux production per depth section 

```{r Flux}
#flux density from middle layer to upper layer umol/m^2/s^1
GWdata$F1 <- -GWdata$deriv1 * GWdata$D 

#flux density from bottom layer to middle layer 
GWdata$F2 <- -GWdata$deriv2 * GWdata$D

#flux for depth section for loop umol/m^3/hr^-1
GWdata$fluxprod <- numeric(n)
for (i in 1:n){
    if (GWdata$F2[i] > GWdata$F1[i]){
      GWdata$fluxprod[i] <- (GWdata$F2[i] - GWdata$F1[i])/0.2 *3600
    }else { 
      GWdata$fluxprod[i] <- (GWdata$F1[i]-GWdata$F2[i])/0.2 *3600
    }
}

#crude plot showing flux density across depth (** note: different locaitons for each gas well)
plot(GWdata$depth, GWdata$fluxprod, 
     xlab = 'depth (cm)',
     ylim = c(0,100000),
     ylab = expression('CO'[2 ]~ 'produced'~ mu ~ mol~ m^-3 ~ hr^-1),
     col = "blue", 
     pch = 19)

```


```{r}
library(xlsx)
library(rJava)
write.xlsx(GWdata, "C:\\Users\\14014\\Desktop\\UH_Soil_Warming\\Gas_wells\\R_output\\2019_10_18_data.xlsx")
```


